---
title: "Homework 1"
author: "STUDENT NAME"
date: "DIFFERENT QUARTER"
output: html_document
---
```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction to Swiss Dataset

The dataset contains information from 47 French-speaking provinces of Switzerland in 1888. The variables include fertility, males in agriculture, high army examination scores, education, percent Catholic, and infant mortality.

```{r, echo=FALSE, warning=FALSE}
library(pander)
pander(head(swiss, 5), style = "rmarkdown", split.tables = 120, caption = "5 Example Provences in Swiss Dataset")

```

### Explore Fertility Data

```{r, echo=FALSE}
m <- mean(swiss$Fertility)
s <- sd(swiss$Fertility)

```
The mean fertility of the Swiss provinces were `r round(m, 2)` with a standard deviation of `r round(s, 2)`.

```{r, echo=FALSE}
hist(swiss$Fertility,
     xlab = "Fertility (Ig)",
     main = "Swiss Province Fertility (1888)",
     border ="pink", 
     col ="maroon")
```

### Further Data Analysis
```{r, echo=FALSE}
mIM <- mean(swiss$Infant.Mortality)
mC <- mean(swiss$Catholic)
mFC <- mean(swiss$Fertility[swiss$Catholic>40])
mFNC <- mean(swiss$Fertility[swiss$Catholic<40])

```
From the first plot below, we can see that there may be a slight correlation between fertility and infant mortality with a greater percentage of children living to be one years old corresponding with provinces with higher fertility rates. The second plot indicates provinces that are less Catholic have a slightly lower average fertility.The mean percentage of Catholic population the Swiss provinces is `r round(mC, 2)`. The mean fertility for provinces that have less than `r round(mC, 2)`% Catholic population is `r round(mFNC, 2)` and the mean fertility for provinces that have more than `r round(mC, 2)`% Catholic population is `r round(mFC, 2)`.

```{r, echo=FALSE}
par(mfrow=c(1,2)) 
plot(swiss$Fertility ~ swiss$Infant.Mortality, 
     ylab = "Fertility (Ig)",
     xlab = "Infant Mortality (%)",
     main = "Fertility and Infant Mortality",
     pch = 18,
     col = "maroon")
plot(swiss$Fertility ~ swiss$Catholic, 
     ylab = "Fertility (Ig)",
     xlab = "Catholic (%)",
     main = "Fertility and Catholicism",
     pch = 18,
     col = "maroon")

```

